---
title: "Assignment 2"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```


# Question:

- Task 1: Read data using a relative path
```{r}
library(tidyverse)
college_data <- read_csv("../collegedata.csv")
```

I used read_csv() from the tidyverse to read the collegedata.csv file. I used a relative path because the data file is in the same folder as my Quarto file.

- Task 2:Determine rows and columns
```{r}
college_data %>%
  summarise(rows = n())

college_data %>%
  summarise(columns = ncol(pick(everything())))
```
I used summarise() and n() to determine the number of rows in the dataset. I also used pick(everything()) with ncol() to determine the number of columns. The dataset contains 7,175 rows and 12 columns.

- Task 3: Use tidyverse to select INSTNM, SAT_AVG, and ADM_RATE
```{r}
college_selected <- college_data %>%
  select(INSTNM, SAT_AVG, ADM_RATE)

college_selected
```
I used the select() function from the tidyverse to keep only the INSTNM, SAT_AVG, and ADM_RATE variables from the dataset. The resulting dataset contains only these three columns.

- Task 4: Create a scatterplot for SAT_AVG and ADM_RATE (Remove NA)
```{r}
college_selected %>%
  drop_na(SAT_AVG, ADM_RATE) %>%
  ggplot(aes(x = SAT_AVG, y = ADM_RATE)) + 
  geom_point() + 
  labs(
    x = "Average SAT Score",
    y = "Admission Rate",
    title = "Average SAT Score vs. Admission Rate"
  )
```
I first removed rows with missing values for SAT_AVG or ADM_RATE using drop_na(). I then used ggplot() and geom_point to create a scatterplot, with average SAT score on the x-axis and admission rate on the y-axis. The plot allows us to see the relationship between these two variables.