---
title: "Assignment 2"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```


# Question:

- Task 1: Read data using a relative path
```{r}
library(tidyverse)
college_data <- read_csv("../collegedata.csv")
```

I used read_csv() from the tidyverse to read the collegedata.csv file. I used a relative path because the data file is in the same folder as my Quarto file.

- Task 2:Determine rows and columns
```{r}
college_data %>%
  summarise(rows = n())

college_data %>%
  summarise(columns = ncol(pick(everything())))
```
I used summarise() and n() to determine the number of rows in the dataset. I also used pick(everything()) with ncol() to determine the number of columns. The dataset contains 7,175 rows and 12 columns.

- Task 3: Use tidyverse to select INSTNM, SAT_AVG, and ADM_RATE

- Task 4: Create a scatterplot for SAT_AVG and ADM_RATE (Remove NA)

## Submission

- After completing **each question**, save your Quarto file, render it, then **stage and commit** your changes.

- Each commit should:

  - Represent **one logical change**.
  - Have a **short, descriptive** commit message.
  - Be written in the **present tense** (imperative mood).

  **Examples:**

  **Bad:** `Fix bug in the project` *Too vague—it does not explain which bug was fixed.*

  **Bad:** `Fixed crash on login screen when password is empty` *Uses the past tense.*

  **Good:** `Fix crash on login screen when password is empty`

- Include well-commented code and clear explanations for each solution.

- Make **at least one meaningful commit per question/part** (a minimum of **4 meaningful commits**).

- **Push your work to your** GitHub\*\*. ane post the link on the text bar of the Canvas assignment.

- Assignments that are **not pushed to GitHub** will receive **no more than 60% of the available points**.

- Submit the rendered HTML file to **Canvas**. Assignments that are **not submitted to Canvas** will also receive **no more than 60% of the available points**.

- If you forget to submit your work to **GitHub** or **Canvas**, you have **up to three (3) days** after the due date to correct the submission. A **10% late penalty** will be applied to the assignment. **After three days, no submissions or corrections will be accepted.**

- **No corrections, re-submissions, or missing submissions will be accepted after the assignment has been graded**, including forgotten GitHub or Canvas submissions.

**Good luck!**
