---
title: "Homework Tidyverse"
subtitle: "Salme Adeele Hollas"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.height = 3,
  fig.width = 6,
  fig.align = "center"
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
#| warning: false
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, gapminder, broom)
```

# Problem 1:

Copy paste and run the tribble given below.

```{r}
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
) -> dt
```

## 1_a:

```{r}
# Calculate the mean of each column using map()
map(dt, mean)
```
I used the map() function to apply the mean() function to each column of dt. This gives the mean value for each variable in the data table.

## 1_b:

```{r}
# Calculate the standard deviation of each column using map()
map(dt, sd)
```
I used map() function to apply the sd() function to each column of dt. This gives the standard deviation for each variable.

## 1_c:

```{r}
# Calculate the square root of every value using map()
map(dt, sqrt)
```
I used map() to apply the sqrt() function to every value in each column of dt.

## 1_d:

```{r}
summary(dt)
```
I used th esummary() function to calculate the minimum, 1st quartile, median, mean, 3rd quartile, and maximum for each column of dt.

# Problem 2:

Write a function that uses a for loop that, for each iteration, randomly draws 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp()`) and calculates its "mean". It should do this 10,000 times. Choose an appropriate plot to plot the distribution of "means".

- Solution: We want to fill a vector with values. We should create this vector beforehand using the `vector()` function.

- We want to create a vector of size 10 and we want its values to be numeric.

- Then we want to randomly draw 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp(5, rate = 1)`) and calculates its "mean"

- Now we want to create a vector that has length of 10 and its values are mean of this randomly selected exponential functions

- Now we want to draw the histogram. But we need to have a data frame. Let us define a data frame.

- Now we can draw a histogram

  - **Now, let's solve the problem all in one step**

## 2_a:

- Repeat part 1 by using a `map_*()` function.

## 2_b:

- Repeat part 1 by using the `replicate()` function.

## 2_c:

- Use a another for loop that will print out plots for sample sizes of 5, 10, and 20 observations (instead of just 5).

# Problem 3:

- Use and show R coding to calculate the "standard deviation" for each variable of the data table `mtcars` using the “Special For Loop Method”.

**Note:**

- There are two methods. Please see a quick explanation for both methods in below:

  - Method 1:

```{r}
  multi_return <- function() {
    my_list <- list("color" = "red", "size" = 20, "shape" = "round")
    return(my_list) 
  }
```

```{r}
a <- multi_return()
a$color

a<- multi_return()
a$shape

a<- multi_return()
a$size
```

- Method 2. The "Special For Loop Method"

```{r}
#| eval: false
output <- vector("double", ncol(df))  # 1. output (use appropriate data frame instead of df)
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output
```

- **Solution:** coding to calculate the "standard deviation" for each variable

# Problem 4:

Generate a sequence of dates from **January 1, 2015** to **December 31, 2025**, spaced by every **two months**. Extract the **year**, **quarter**, and **ISO week number** for each date.

# Problem 5

Given the following dates, compute the difference in **months** and **weeks** between each consecutive pair.

``` r
sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")
```

# Problem 6

Given a list of mixed date formats, use `map()` and `possibly()` from `purrr` to safely convert them to **Date** format and extract the **month name**.

``` r
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")
```
