---
title: "Homework Tidyverse"
subtitle: "Salme Adeele Hollas"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.height = 3,
  fig.width = 6,
  fig.align = "center"
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
#| warning: false
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, gapminder, broom)
```

# Problem 1:

Copy paste and run the tribble given below.

```{r}
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
) -> dt
```

## 1_a:

```{r}
# Calculate the mean of each column using map()
map(dt, mean)
```

I used the map() function to apply the mean() function to each column of dt. This gives the mean for each variable.

## 1_b:

```{r}
# Calculate the standard deviation of each column using map()
map(dt, sd)
```

I used the map() function to apply the sd() function to each column of dt. This gives the standard deviation for each variable.

## 1_c:

```{r}
# Calculate the square root of every value using map()
map(dt, sqrt)
```

I used the map() to apply the sqrt() function to every value in each column of dt.

## 1_d:

```{r}
summary(dt)
```

I used the summary() function to calculate the minimum, 1st quartile, median, mean, 3rd quartile, and maximum for each column of dt.

# Problem 2:

Write a function that uses a for loop that, for each iteration, randomly draws 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp()`) and calculates its "mean". It should do this 10,000 times. Choose an appropriate plot to plot the distribution of "means".

- Solution: We want to fill a vector with values. We should create this vector beforehand using the `vector()` function.

- We want to create a vector of size 10 and we want its values to be numeric.

- Then we want to randomly draw 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp(5, rate = 1)`) and calculates its "mean"

- Now we want to create a vector that has length of 10 and its values are mean of this randomly selected exponential functions

- Now we want to draw the histogram. But we need to have a data frame. Let us define a data frame.

- Now we can draw a histogram

  - **Now, let's solve the problem all in one step**

```{r}
# Function to generate 10,000 sample means
calculate_means <- function() {
  means <- vector("double", 10000)
  
  for (i in 1:10000) {
    sample <- rexp(5, rate = 1)
    means[i] <- mean(sample)
  }
  means
}
```

```{r}
means <- calculate_means()
head(means)
```

```{r}
# Create a data frame and plot the distribution of the sample means
means_df <- tibble(means = means)

ggplot(means_df, aes(x=means)) +
  geom_histogram(binwidth = 0.05) +
  labs(
    x= "Sample Mean",
    y = "Count",
    title = "Distribution of Sample Means"
  )
```

I stored the 10,000 sample means in a data frame and used a histogram to show their distribution.

## 2_a:

```{r}
# Generate 10,000 sample means using map()
means_map <- map_dbl(1:10000, ~ mean(rexp(5, rate = 1)))
head(means_map)
```

```{r}
# Plot the distribution of the sample means
means_map_df <- tibble(means = means_map)
ggplot(means_map_df, aes(x=means)) +
  geom_histogram(binwidth = 0.05) + 
  labs(
    x = "Sample Mean",
    y = "Count",
    title = "Distribution of Sample Means Using map()"
  )
```

I used map_dbl() to generate 10,000 samples and calculate the mean of each sample. The histogram shows the distribution of these sample means.

## 2_b:

```{r}
# Generate 10,000 sample means using replicate()
means_replicate <- replicate(10000, mean(rexp(5, rate = 1)))
head(means_replicate)
```

```{r}
means_replicate_df <- tibble(means = means_replicate)
ggplot(means_replicate_df, aes(x = means)) +
  geom_histogram(binwidth = 0.05) +
  labs(
    x = "Sample Mean",
    y = "Count",
    title = "Distribution of Sample Means Using replicate()"
  )
```

I used replicate() to repeat the sampling and mean calculation 10,000 times. The histogram shows the distribution of these sample means.

## 2_c:

```{r}
# Generate sample means for sample sizes 5, 10, and 20
sample_sizes <- c(5, 10, 20)

for (n in sample_sizes) {
  means <- vector("double", 10000)
  for (i in 1:10000) {
    sample <- rexp(n, rate =1)
    means[i] <- mean(sample)
  }
  
  means_df <- tibble(means = means)
  
  print(
    ggplot(means_df, aes(x=means)) +
      geom_histogram(binwidth = 0.05) +
      labs(
        x="Sample Mean",
        y= "Count",
        title = paste("Sample Size =", n)
      )
  )
}
```

I used a for loop to generate 10,000 sample means for sample sizes of 5, 10, and 20. The histograms show that the distribution becomes narrower as the sample size increases.

# Problem 3:

```{r}
# Calculate the standard deviation of each variable in mtcars
output <- vector("double", ncol(mtcars))

for (i in seq_along(mtcars)) {
  output[[i]] <- sd(mtcars[[i]])
}

names(output) <- names(mtcars)
output
```

I used the special for loop method to calculate the standard deviation of each variable in mtcars. The loop goes through each column and applies the sd() function.

# Problem 4:

```{r}
# Generate dates every 2 months from 2015 to 2025
dates <- tibble(
  date = seq(
    as.Date("2015-01-01"),
    as.Date("2025-12-31"),
    by = "2 months"
  )
) %>%
  mutate(
    year = year(date),
    quarter = quarter(date),
    iso_week = isoweek(date)
  )

dates
```

I created a sequence of dates every two months and used lubridate functions to extract the year, quarter, and ISO week number for each date.

# Problem 5

```{r}
sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")

# Convert the date strings to Date format
dates <- ymd(sample_dates)

# Calculate the difference between consecutive dates
date_differences <- tibble(
  from = dates[-length(dates)],
  to = dates[-1],
  months = time_length(interval(from, to), "months"),
  weeks = time_length(interval(from, to), "weeks")
)

date_differences
```

I converted the date strings to Date format and calculated the difference in months and weeks between each consecutive pair of dates.

# Problem 6

```{r}
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")

# Safely convert different date formats to Date
safe_date <- possibly(
  function(x) {
    as.Date(parse_date_time(x, orders = c("ymd", "Y/m/d", "d-b-Y")))
  },
  otherwise = as.Date(NA)
)

# Convert the date strings
months <- map_chr(
  dates,
  ~ as.character(month(.x, label = TRUE, abbr = FALSE))
)

# Extract the month name
results <- tibble(
  date_string = unlist(date_strings),
  date = as.Date(unlist(dates)),
  month = unlist(months)
)


results
```

I used parse_date_time() to handle the different date formats and possibly() so that the invalid date would return NA instead of causing an error. I then used map() to convert the dates and extract the month names.
